Recurrence of Geodesic Flow on the Necker Cube Surface

### 2017-08-09 paper.tex
070917

\newtheorem{thm}{Theorem}[subsection]
\newtheorem{Def}{Definition}[subsection]
\newtheorem*{thm*}{Theorem}
\newtheorem*{def*}{Definition}
\newtheorem{lem}{Lemma}[subsection]
\newtheorem*{rem}{Remark}

@@

\subsection{Some Background}
This paper assumes that the reader is familiar with fundamental concepts in algebraic topology, such as covering space theory, and the study of Veech surfaces and their associated Veech groups. In order to prove this theorem, we take a number of covers of a surface isometric to $\mathbf S$. Our goal then is to take a branched cover of the surface as four "copies" of that surface, and rotate these copies so as to align the geodesic flow and identify edges solely by translation. The quotient space obtained by modding out this surface by its translational symmetries is translation and Veech.

@@

\subsection{A Path-Isometric Map From $\mathbf S$ to $\mathbf U$}
A path-isometric map from $\mathbf{S}$ to $\mathbf{P}$ induces an isometry from the Necker cube surface to $\mathbf{U}$. The path isometry $\Psi:\mathbf{S}\rightarrow\mathbf{P}$ is a piece-wise, well-defined map that is given as a linear transformation in which a square labeled $\mathbf{A}$, $\mathbf{B}$, or $\mathbf{C}$ is mapped to its corresponding square on $\mathbf{P}$. $\Psi$ is  essentially a transformation that cuts along the edges parallel with the z-axis. (See figure $\ref{fig:cut}$)

@@

\begin{Def}
The translation surface $\mathbf{M}$ is given as the quotient  $\tilde{\mathbf{U}}/\mathbf{T}^{m,n}$.
\end{Def}

@@

\subsection{$\mathbf M$ as a Veech Surface}
By a series of cutting and gluing operations, $\mathbf M$ has this symmetric representation as a square-tiled staircase (figure $\ref{fig:staircase}$). The vertices of cone angle $2\pi$ can be found on the centers of the squares in the figure below where the black lines intersect. 
\begin{figure}[H]
\centering
\includesvg[width=4in]{mtildastaircase}
\caption{$\mathbf{M}$ as a square-tiled staircase with cone singularities identified.}
\label{fig:staircase}
\end{figure}
As a result of Gutkin-Judge [citation], $\mathbf M$ is a Veech surface whose Veech group is commensurable to $SL(2,\mathbb Z)$. In order to determine which elements of orientation preserving affine diffeomorphisms, denoted Aff$^{+}(\mathbf M)$, have derivatives that belong to the Veech group, V($\mathbf M$), it is necessary to study $\mathbf M$'s various symmetries and cylinder decompositions.
\vspace{0.1in}

@@

One group contained in $V(\mathbf M)$ is SO$(2,\mathbb{Z})$, the group generated by rotations of order 4. Even though the other elements in $\mathbb{D}_4$ are derivatives of automorphisms of $\mathbf M$, we will only consider those of determinant +1. We want the corresponding affine map of these rotation matrices to obey the convention that we \emph{fix} the second square on the bottom with label I. These maps are unique only up to translation and as such assigning this convention makes it easier to keep track of where the edges of $\mathbf M$ end up afterwards. The diffeomorphism with 90 degree counter-clockwise rotation fixed about that point at the center of that square is denoted $\mathbf{R}$. Its derivative, $\mathbf{R}'$, is the matrix $\left[ \hspace{1mm} \begin{matrix}
							0 & -1\\
							1 & 0
						\end{matrix}\hspace{1mm}\right]$.

@@

The translational symmetries have no unique representation as an element of V($\mathbf M$), as their derivatives are represented as the identity. But they do have an important effect on the surface. Shifting the bottom left-most square to the right and reconstructing the rest by pairing edges is an order 2 homeomorphism denoted $\mathbf{H}$, for horizontal. Likewise, $\mathbf{V}$ is an order 2 homeomorphism that shifts that square upwards. Observe that together they satisfy the property $(\mathbf{H}\mathbf{V})^{6}=(\mathbf{V}\mathbf{H})^{6}=id$. This essentially moving the squares up the staircase until they come back to where they started. The group of translation deck transformations of the surface is isomorphic to a permutation group of 12 elements (each square) generated by these two translations that satisfy the relation. It is the group $\mathbb{D}_{12}$ generated by reflections about two adjacent vertices of a 12-gon. Thus it is possible to translate a square to any other on the surface as a composition of these two translations.

@@

Both sets of cylinders are indexed $0\leq i, j \leq 5$. Every cylinder is composed of two unit squares with edges identified by translation. We start counting the cylinders from the bottom-most cylinder and increment the index as the cylinders fill the surface going ``up" the staircase. $\small{\mathcal{E}}_0$ is the first cylinder at the bottom whose identifications are \em{not} \normalfont the solid triangles. $\small{\mathcal{E}}_5$ is the cylinder whose edges are paired by the solid triangles. $\Delta_{i}$ has the more standard representation starting from the bottom-most horizontal cylinder.

We define the \emph{inverse modulus} of a cylinder to be the ratio of a cylinders circumference to its width, $\mu^{-1}=\frac{c}{h}$. A \emph{Dehn-twist} of a cylinder is an automorphism of the cylinder of the form  $(x,y)\mapsto(x,y+\mu^{-1} x \mod{h})$, where $x$ and $y$ are points on the cylinder such that x varies along the its circumference. According to [need to cite something], a translation surface whose finite cylinder decompositions in orthogonal directions form a set of inverse moduli. The smallest (rational) common multiple of the set of all these values match up to give a global affine diffeomorphisms of the surface, whose derivatives have the matrix representations
\begin{equation*}
\left[ \hspace{1mm} \begin{matrix}
				1 & \pm \lambda\\
				0 & 1
			\end{matrix}\hspace{1mm}\right] \text{ and }
			\left[ \hspace{1mm} \begin{matrix}
							1 & 0\\
							\pm \lambda & 1
						\end{matrix}\hspace{1mm}\right],
\end{equation*}
such that $\lambda$ is that lowest common multiple.

Let $\mu^{-1}_{\Delta_i},\mu^{-1}_{\mathcal{E}_j}$ denote the inverse moduli of the cylinders. The cylinders $\Delta_{i}$ and $\small{\mathcal{E}}_j$, as i and j vary, all have the same inverse moduli since they are all defined as two unit squares with identifications made by translation along the cylinders circumference. Therefore, $\lambda=$ l.c.m.($\mu^{-1}_{\Delta_i}\cup\mu^{-1}_{\mathcal{E}_j}$) $=2$.

@@

It is now possible to generate the \emph{group of symmetries} of $\textbf{M}$ as Deck transformations of the surface. We call this group $\mathbb X$ and say that

@@

The group of derivatives of $\mathbb X$ is $\mathbb X'$, and it is generated by elements of V$(\textbf{M})$. It is not apparent whether or not $\mathbb{X}=$Aff$^+(\mathbf{M})$ or if $\mathbb X'=$ V$(\mathbf{M})$.

\subsection{Action of $\mathbb X$ on the Homology of $\mathbf M$}
The group $\mathbb X$ consists of the set of Deck transformations of $\mathbf{M}$, whose induced isomorphisms of $\pi_1(\mathbf{M})$ acts on the absolute homology of $\mathbf{M}$ and extends to an action on the homology of $\mathbf{M}$'s $\mathbb{Z}^2$-cover, $\tilde{\mathbf{U}}$. $\mathbf{M}$'s cylinder decomposition gives us a natural set of homology classes that serve as a basis for $H_1(\mathbf{M},\mathbb Q)$. For a total of 12 cylinders, there are 12 core curves:

\begin{figure}[H]
\includegraphics[width=3in]{homologyclass.png}
\centering
\caption{Cylinder core curves with u,v, and z homology classes that determines the $\mathbb{Z}^2$-cover.}
\end{figure}

@@

We will denote the group of isomorphisms of $\mathbf{M}$'s fundamental group we obtain from $\mathbb{X}$ as $\mathbb X_*$. The same notation will be used for the isomorphisms themselves. We denote the \emph{set of cylinder core curves} as $\Gamma$.

@@

The 12-gon was mentioned in the previous section under the discussion of $\mathbf{M}$'s translational symmetries. $\mathbb X_*$ acting on absolute homology is better represented as a symmetric object.

@@

This graph's representation also reflects how two core curves are incident with only two others. An element of $\mathbb X_*$ either permutes these curves or, in the case where $\mathbf A_*$ or $\mathbf B_*$ generates the element, adds/subtracts black vertices to/from the red ones. Recall that in the previous section we had certain conventions when it came to the elements $\mathbf{R}$, $\mathbf{H}$, and $\mathbf{V}$. This was so that their induced isomorphisms would \emph{fix} one or more vertices as they act on homology. But because of the translational symmetries, it is possible to rotate about the center of any square and essentially rotating the entire 12-gon with $\mathbf{H}_*$ and $\mathbf{V}_*$.

@@

The two preserve the signs of the core curves. It is to be expected as they are induced isomorphisms of translation maps. They are given as the following permutations of $\Gamma$:
\space
\begin{align*}
\mathbf{H}_* &= (\gamma_0)(\gamma_1\hspace{1mm}\gamma_{11})(\gamma_2\hspace{1mm}\gamma_{10})(\gamma_3\hspace{1mm}\gamma_9)(\gamma_4\hspace{1mm}\gamma_8)(\gamma_5\hspace{1mm}\gamma_7)(\gamma_6)\\
\mathbf{V}_* &=(\gamma_{11})(\gamma_0\hspace{1mm}\gamma_{10})(\gamma_1\hspace{1mm}\gamma_{9})(\gamma_2\hspace{1mm}\gamma_8)(\gamma_3\hspace{1mm}\gamma_7)(\gamma_4\hspace{1mm}\gamma_6)(\gamma_5)
\end{align*}

Rotation about the center of a square will not only change signs of the core curves, but permute elements in such a way that 4 consecutive rotations of $\mathbf{M}$ is an identity on $\Gamma$.

\begin{minipage}{0.4\textwidth}
\begin{figure}[H]
\hspace{0}\includegraphics[width=1.7in]{12gonR.pdf}
\end{figure}
\end{minipage}
\begin{minipage}{0.55\textwidth}
\vspace{0in}
\begin{itemize}
\item[\textbf{\emph{$\mathbf{R}_*$}}] This action reflects about the line and maps one vertex to another. No vertex is fixed. However it has the additional effect of changing the sign of a curve only if it was originally black (or odd index). The pattern is peculiar, but $\mathbf{R}_*$ is still order four.
\end{itemize}
\end{minipage}

\vspace{2mm}
We can see the effect that it has on homology clearer as a permutation of $\Gamma$:
\begin{align*}
\mathbf{R}_*= \hspace{1mm} &(\gamma_{0}\hspace{1mm}\gamma_{1}\hspace{1mm}\text{-}\gamma_{0}\hspace{1mm}\text{-}\gamma_{1})
(\gamma_{2}\hspace{1mm}\gamma_{11}\hspace{1mm}\text{-}\gamma_{2}\hspace{1mm}\text{-}\gamma_{11})
(\gamma_{3}\hspace{1mm}\text{-}\gamma_{10}\hspace{1mm}\text{-}\gamma_{3}\hspace{1mm}\gamma_{10})\\
&(\gamma_{4}\hspace{1mm}\gamma_{9}\hspace{1mm}\text{-}\gamma_{4}\hspace{1mm}\text{-}\gamma_{9})
(\gamma_{5}\hspace{1mm}\text{-}\gamma_{8}\hspace{1mm}\text{-}\gamma_{5}\hspace{1mm}\gamma_{8})
(\gamma_{6}\hspace{1mm}\gamma_{7}\hspace{1mm}\text{-}\gamma_{6}\hspace{1mm}\text{-}\gamma_{7}).
\end{align*}

\compav{Ended here August 3.}

Regardless, $u$ and $v$ are the most important homology classes that we use to determine whether or not a lifted path properly closes on the four-fold cover.  Let $\alpha:[0,1]\rightarrow\mathbf M$ be a curve on the surface such that $\alpha(0)=\alpha(1)=x_0\in\mathbf{M}$. Take $f:\tilde{\mathbf U}\rightarrow\mathbf M$ to be the covering map that takes every point on $\tilde{\mathbf U}$ to its quotient in $\mathbf{M}$. We will fix a point $\tilde{x}_0\in\tilde{\mathbf U}$ such that the lifted path $\tilde\alpha(0)=\tilde{x}_0$ under the function $f$. $\alpha$ is a closed path in $\pi_1(\mathbf M,x_0)$. We denoted the class of abelianized paths $\alpha$ as [$\alpha$]$\in H_1(\tilde{\mathbf{M}},\mathbb Z)$, and define the function,

\begin{align}
i:H_1(\mathbf{M},\mathbb Z)\times H_1(\mathbf{M},\mathbb Z)\rightarrow \mathbb Z,
\end{align}

to return the intersection number of two homology classes. Out of convention, we consider a positive intersection to adhere to the right-hand rule and take the crossings to be relative to $u$ and $v$. With that being said, the signed homology classes of gamma curves are

@@

The $\mathbb{Z}^2$ cover, $\tilde{\mathbf{U}}$, is associated to the kernel of the homomorphism
\begin{align}
\Omega_{u,v}:\pi_1(\mathbf{M},x_0)\rightarrow \mathbb{Z}^2, \text{ such that } \alpha\mapsto(i(u,[\alpha]),i(v,[\alpha])).
\end{align}
Suppose that $h:\mathbf{M}\rightarrow\mathbf{M}$ is an automorphism of the surface such that dir $h\in SL(\mathbf M)$. $h$ induces the isomorphism, $h_*:\pi_1(\mathbf{M},x_0)\rightarrow \pi_1(\mathbf{M},x_0)$. Likewise, $f$ induces the inclusion map $f_*:\pi_1(\tilde{\mathbf{U}},\tilde x_0)\hookrightarrow\pi_1(\mathbf{M},x_0)$. If $\pi_1(\tilde{\mathbf{U}},\tilde x_0)$ is associated with the Ker $\Omega_{u,v}$, then it must be true that the kernel of such a homomorphism is preserved under $h_*$ and there exists a map from  $\pi_1(\mathbf M, x_0)$ to $\pi_1(\tilde{\mathbf{U}},\tilde x_0)$.

The actions of $h_*$ are dual to the actions of dir $h$, denoted $h'$, on the sets of odd-odd and even-odd $\mathbb{Z}^2$ vectors, $\mathcal{O}$ and $\mathcal{E}$. In particular, the action of $\mathbb X$ on $\mathcal{O}$ and $\mathcal{E}$ is self-contained and well-defined. This is vital as what we want to show is that $h'(\alpha')$, where $\alpha'$ is a constant and $\alpha$ is a closed geodesic, is dual to $h_*([\alpha])$. In which case, it should also be true that $h_*(\alpha)\mapsto(i(u,h_*([\alpha])),i(v,h_*([\alpha])))$, under $i$. And since intersection number is a non-degenerate bilinear form,

\begin{align*}
\alpha\mapsto(i(h_*^{-1}(u),[\alpha]),i(h_*^{-1}(v),[\alpha])).
\end{align*}

This amounts to finding some such path whose lift belongs to the fundamental group of $\tilde{\mathbf U}$, noting the effect $h_*$ has on $\mathbf M$'s homology, and finding the new intersection number. Let $\beta$ be such a path. Suppose there is some sequence of affine maps $h_1,...,h_n\in$Aff$^+(\mathbf{M})$ such that $h_1' \circ ... \circ h_n'(\alpha')=\beta'$. We will show the following is true:

\begin{align*}
h_1 \circ ... \circ h_n(\mathbf{M})=\mathbf{M}'&\iff\\
h_{*_1} \circ ... \circ h_{*_n}([\beta])=[\alpha] &\iff
\end{align*}
%\subsection{$\tilde{\mathbf{M}}$ as a Veech Surface}
%We use basic properties of $\tilde{\mathbf{M}}$ to show that it is Veech, and its Veech group is a lattice in $SL_2(\mathbb{Z})$. Recall that translation surfaces are equipped with charts, $\omega$, of maps from the surface to $\mathbb{C}$. The Veech group of the surface $\tilde{\mathbf{M}}$ is given as the image of its stabilizer in the projective group $PSL_2(\mathbb{R})$.
%
%\begin{Def}
%Since $\tilde{\mathbf{M}}$ is a translation surface with four singularities, it is a Veech surface that locally behaves like a torus away from these cone singularities. $\tilde{\mathbf{M}}$ has a \textbf{Veech group}, denoted $\Gamma$.
%\end{Def}
%
%We use the following theorem of Gutkin and Judge to show that $\Gamma$ is lattice (and arithemtic) in $SL_2(\mathbb{Z})$:
%
%\begin{thm}(Gutkin-Judge) A translation surface $(S,\omega)$ is squaretiled
%if and only if its Veech group $V(S,\omega)$ shares a finite-index subgroup
%with $SL_2(\mathbb{Z}).$
%\end{thm}
%
%\begin{thm}The following are equivalent:
%\begin{enumerate}[label=(\roman*)]
%\item $\tilde{\mathbf{M}}$ is tiled by squares.
%\item $\Gamma$ is lattice in $SL_2(\mathbb{Z})$.
%\end{enumerate}
%\begin{proof}
%figure $\ref{fig:staircase}$ explicitly demonstrates how $\tilde{\mathbf{M}}$ is tiled by squares. The proof then follows as an obvious result of the previous theorem.
%\end{proof}
%\end{thm}

>>>>

\newtheorem{thm}{Theorem}[section]
\newtheorem{Def}{Definition}[section]
\newtheorem*{thm*}{Theorem}
\newtheorem*{def*}{Definition}
\newtheorem{lem}{Lemma}[section]
\newtheorem*{rem}{Remark}

@@

\subsection{Some Background}
This paper assumes that the reader is familiar with fundamental concepts in algebraic topology, such as covering space theory and the study of Veech surfaces and associated Veech groups. In order to prove this theorem, we take a number of covers of a surface isometric to $\mathbf S$. Our goal then is to take a branched cover of the surface as four "copies" of that surface, and rotate these copies so as to align the geodesic flow and identify edges solely by translation. The quotient space obtained by modding out this surface by its translational symmetries is translation and Veech.

@@

\subsection{A Path Isometry From $\mathbf S$ to $\mathbf U$}
A path-isometric map from $\mathbf{S}$ to $\mathbf{P}$ induces an isometry from the Necker cube surface to $\mathbf{U}$. The path isometry $\Psi:\mathbf{S}\rightarrow\mathbf{P}$ is a piece-wise, well-defined map that is given as a linear transformation in which a square labeled $\mathbf{A}$, $\mathbf{B}$, or $\mathbf{C}$ is mapped to its corresponding square on $\mathbf{P}$. $\Psi$ is  essentially a transformation that cuts along the edges parallel with the z-axis. (See figure $\ref{fig:cut}$)

@@

\begin{Def}
The translation surface $\mathbf{M}$ is given as a quotient of $\tilde{\mathbf{U}}$ equipped with a  holomorphic 1-form, $(\tilde{\mathbf{U}}/\mathbf{T}^{m,n},\omega)$.
\end{Def}

@@

\subsection{$\mathbf M$ as a Veech Surface}
By a series of cutting and gluing operations, $\mathbf M$ has this symmetric representation as a square-tiled staircase (figure $\ref{fig:staircase}$). This is a compact surface with an infinite $\mathbb Z$-cover as the Infinite Staircase. [cite] It shares many of the same symmetries, as well. The vertices of cone angle $2\pi$ can be found on the centers of the squares in the figure below where the black lines intersect. 
\begin{figure}[H]
\centering
\includesvg[width=4in]{mtildastaircase}
\caption{$\mathbf{M}$ as a square-tiled staircase with cone singularities identified.}
\label{fig:staircase}
\end{figure}
As a result of Gutkin-Judge [citation], $\mathbf M$ is a Veech surface whose Veech group is commensurable to $SL(2,\mathbb Z)$. In order to determine which elements of orientation preserving affine diffeomorphisms, denoted Aff$^{+}(\mathbf M)$, have derivatives that belong to the Veech group, V($\mathbf M$), it is necessary to study $\mathbf M$'s various symmetries, as well as its cylinder decomposition.
\vspace{0.1in}

@@

One group contained in $V(\mathbf M)$ is SO$(2,\mathbb{Z})$, the group generated by rotations of the square. We want the corresponding affine map of these rotation matrices to obey the convention that we \emph{fix} the second square on the bottom with label I. These maps are unique only up to a translation of the surface and assigning this convention makes it easier to keep track of where the edges of $\mathbf M$ end up. The 90 degree counter-clockwise rotation fixed about the point in the center of that square is denoted $\mathbf{R}$. Its derivative, $\mathbf{R}'$, is the matrix $\left[ \hspace{1mm} \begin{matrix}
							0 & -1\\
							1 & 0
						\end{matrix}\hspace{1mm}\right]$.

@@

The translational symmetries have no unique representation as an element of V($\mathbf M$), as their derivatives are represented as the identity. But they do have an important effect on the surface. Shifting the bottom left-most square to the right and reconstructing the rest by pairing edges is an order 2 homeomorphism denoted $\mathbf{H}$, for horizontal. Likewise, $\mathbf{V}$ is an order 2 homeomorphism that shifts that square upwards. Observe that together they satisfy the property $(\mathbf{H}\mathbf{V})^{6}=(\mathbf{V}\mathbf{H})^{6}=id$. This is essentially moving the squares up the staircase until they come back to where they started, unlike the the group of diagonal translations of the Infinite Staircase that is isomorphic to $\mathbb Z$. The group Trans($\mathbf M$) is isomorphic to the group generated by reflections of a 12-gon, $\mathbb{D}_{12}$.

@@

Both sets of maximal cylinders are indexed $0\leq i, j \leq 5$. Every cylinder is composed of two unit squares with edges identified by translation. Each cylinder is obtained as a ramified cover of the square torus. We start counting the cylinders from the bottom-most cylinder and increment the index as the cylinders fill the surface going ``up" the staircase. $\small{\mathcal{E}}_0$ is the first cylinder at the bottom whose identifications are \em{not} \normalfont the solid triangles. Travel up the staircase until the laste one, $\small{\mathcal{E}}_5$, the cylinder whose edges are paired by the solid triangles and glue the 12th square with the 1st. $\Delta_{i}$ has the more standard representation starting from the bottom-most horizontal cylinder and going all the way up to the top.

We define the \emph{inverse modulus} of a cylinder to be the ratio of a cylinders circumference to its width, $\mu^{-1}=\frac{c}{h}$. A \emph{Dehn-twist} of a cylinder is an automorphism of the cylinder of the form  $(x,y)\mapsto(x,y+\mu^{-1} x \mod{h})$, where $x$ and $y$ are points on the cylinder such that x varies along the its circumference. According to [need to cite something], a translation surface whose finite cylinder decompositions in orthogonal directions form a set of inverse moduli. The smallest (rational) common multiple of the set of all these values match up to give a global affine diffeomorphisms of the surface, whose derivatives have the matrix representations
\begin{equation*}
\left[ \hspace{1mm} \begin{matrix}
				1 & \pm \lambda\\
				0 & 1
			\end{matrix}\hspace{1mm}\right] \text{ and }
			\left[ \hspace{1mm} \begin{matrix}
							1 & 0\\
							\pm \lambda & 1
						\end{matrix}\hspace{1mm}\right],
\end{equation*}
such that $\lambda$ is that lowest common multiple. [cite]

Let $\mu^{-1}_{\Delta_i},\mu^{-1}_{\mathcal{E}_j}$ denote the inverse moduli of the cylinders. The cylinders $\Delta_{i}$ and $\small{\mathcal{E}}_j$, as i and j vary, all have the same inverse moduli since they are all defined as two unit squares with identifications made by translation. Therefore, $\lambda=$ l.c.m.($\mu^{-1}_{\Delta_i}\cup\mu^{-1}_{\mathcal{E}_j}$) $=2$.

@@

It is now possible to generate the \emph{group of symmetries} of $\textbf{M}$ with the elements of these affine transformations. We call this group $\mathbb X$ and say that

@@

The group of derivatives of $\mathbb X$ is the image of the map D: $\mathbb X\rightarrow SL(2,\mathbb{R})$, denoted $\mathbb X'$. We know that $\mathbb{X}\subset$Aff$^+(\mathbf{M})$ and $\mathbb X'\subset$ V$(\mathbf{M})$. The corresponding elements of $\mathbb X'$ will use the same prime notation.

\subsection{Action of $\mathbb X$ on the Homology of $\mathbf M$}
The elements of $\mathbb X$ induces isomorphisms of $\pi_1(\mathbf{M})$ and extends to an action on the homology of $\mathbf{M}$'s $\mathbf{T}^{m,n}$-cover, $\tilde{\mathbf{U}}$.

$\mathbf{M}$'s cylinder decomposition gives us a natural set of homology classes that serve as a basis for $H_1(\mathbf{M},\mathbb Q)$. For a total of 12 cylinders, there are 12 core curves:

\begin{figure}[H]
\includegraphics[width=3in]{homologyclass.png}
\centering
\caption{Cylinder core curves with u,v, and z homology classes that determines the $\mathbf{T}^{m,n}$-cover.}
\label{fig:homology}
\end{figure}

@@

We will denote the group of isomorphisms of $\mathbf{M}$'s fundamental group we obtain from $\mathbb{X}$ as $\mathbb X_*$. The same notation will be used for the isomorphisms themselves. We denote the \emph{set of cylinder core curves} as $\Gamma\subset H_1(\mathbf{M},\mathbb Q)$.

@@

The 12-gon was mentioned in the previous section under the discussion of $\mathbf{M}$'s translational symmetries. $\mathbb X_*$ acting on absolute homology is better represented as this symmetric object.

@@

This graph's representation also reflects how two core curves are incident with only two others. An element of $\mathbb X_*$ either permutes these curves or, in the case where $\mathbf A_*$ or $\mathbf B_*$ generates the element, adds/subtracts black vertices to/from the red ones. Recall that in the previous section we had certain conventions when it came to the elements $\mathbf{R}$, $\mathbf{H}$, and $\mathbf{V}$. This was so that their induced isomorphisms would \emph{fix} one or more vertices as they act on homology. Because of the translational symmetries, it is possible to rotate or reflect entire 12-gon with $\mathbf{H}_*$ and $\mathbf{V}_*$. These permutations of $\Gamma$ generate a group isometric to $\mathbb{D}_{12}$.

@@

The two preserve the signs of the core curves. It is to be expected as they are induced isomorphisms of translations. They are expressed as:
\space
\begin{align*}
\mathbf{H}_* &= (\gamma_0)(\gamma_1\hspace{1mm}\gamma_{11})(\gamma_2\hspace{1mm}\gamma_{10})(\gamma_3\hspace{1mm}\gamma_9)(\gamma_4\hspace{1mm}\gamma_8)(\gamma_5\hspace{1mm}\gamma_7)(\gamma_6)\\
\mathbf{V}_* &=(\gamma_{11})(\gamma_0\hspace{1mm}\gamma_{10})(\gamma_1\hspace{1mm}\gamma_{9})(\gamma_2\hspace{1mm}\gamma_8)(\gamma_3\hspace{1mm}\gamma_7)(\gamma_4\hspace{1mm}\gamma_6)(\gamma_5)
\end{align*}

Rotation about the center of a square will not only change signs of the core curves, but permute elements in such a way that the rotation element is order four.

\begin{minipage}{0.4\textwidth}
\begin{figure}[H]\includegraphics[width=1.7in]{12gonR.pdf}
\end{figure}
\end{minipage}
\begin{minipage}{0.55\textwidth}
\vspace{0in}
\begin{itemize}
\item[\textbf{\emph{$\mathbf{R}_*$}}] This action reflects the 12-gon about that line and maps one vertex to another. No vertex is fixed. However it has the additional effect of changing the sign of a curve only if it was originally black (or odd index).
\end{itemize}
\end{minipage}

\vspace{2mm}
We can see the effect that it has on homology clearer as a permutation group on $\Gamma$:
\begin{align*}
\mathbf{R}_*= \hspace{1mm} &(\gamma_{0}\hspace{1mm}\gamma_{1}\hspace{1mm}\text{-}\gamma_{0}\hspace{1mm}\text{-}\gamma_{1})
(\gamma_{2}\hspace{1mm}\gamma_{11}\hspace{1mm}\text{-}\gamma_{2}\hspace{1mm}\text{-}\gamma_{11})
(\gamma_{3}\hspace{1mm}\text{-}\gamma_{10}\hspace{1mm}\text{-}\gamma_{3}\hspace{1mm}\gamma_{10})\\
&(\gamma_{4}\hspace{1mm}\gamma_{9}\hspace{1mm}\text{-}\gamma_{4}\hspace{1mm}\text{-}\gamma_{9})
(\gamma_{5}\hspace{1mm}\text{-}\gamma_{8}\hspace{1mm}\text{-}\gamma_{5}\hspace{1mm}\gamma_{8})
(\gamma_{6}\hspace{1mm}\gamma_{7}\hspace{1mm}\text{-}\gamma_{6}\hspace{1mm}\text{-}\gamma_{7}).
\end{align*}

\compav{TODO: Make the parabolic 12-gon image.}

\begin{figure}[H]
\centering
\includesvg{12gonA}
\end{figure}

\begin{figure}[H]
\centering
\includesvg{12gonABexample}
\end{figure}
\compav{Ended here August 3.}

\subsection{$\mathbf{T}^{m,n}$-Cover of $\mathbf M$}
When defining $\mathbf{M}$ as a quotient over the action of $\mathbf{T}^{m,n}$, the result was a finite-type translation surface whose cover indexes these four "cut-open tori" by the integer pair $(m,n)$. For this reason it is easy to consider $\tilde{\mathbf{U}}$ to be a kind of $\mathbb{Z}^2$-esque covering space of $\mathbf{M}$.

\begin{lem}
The family of translational symmetries $\mathbf{T}^{m,n}$ is isomorphic to $\mathbb Z^2$.
\begin{proof}
$\mathbf{T}^{m,n}$ is induced by rotational and translational symmetries of the fundamental domain of $\tilde{\mathbf{U}}$, $\mathbf P \times \mathbb{Z}/4\mathbb{Z}$. It's clear that the rotations themselves are bijective maps. By definition, $\mathbf{T}_0^{m,n}=(x+2m,y+2n;j)$. As a group action, observe that the points $(x,y;j)$ are stabilized. The orbit of this triple is the set of infinitely many points in its domain parametrized by integers $m,n$. The obvious isomorphism identifies an element of $\mathbb{Z}^2$ with $(m,n)$. The orbit of any point is itself given as $\mathbf{T}_0^{\mathbb Z^2}$. Therefore, the induced map on $\tilde{\mathbf{U}}$, $\mathbf{T}^{m,n}$, is isomorphic to $\mathbb Z^2$ as well.
\end{proof}
\end{lem}

From figure $\ref{fig:mtilda}$ we construct homology classes $u,v$ according to the outer edges of the tori. In addition, there is a class $z$ that doesn't play a role in determining the cover $\tilde{\mathbf U}$. This is because $z$ is the set of core curves that lie on cylinders that go from one plane to the next. Figure $\ref{fig:homology}$ labels them on the staircase model.

@@

Let i be a bilinear, non-degenerate form given as

\begin{align}
i:H_1(\mathbf{M},\mathbb Q)\times H_1(\mathbf{M},\mathbb Q)\rightarrow \mathbb Z.
\end{align}

Let $\beta\in \pi_1(\mathbf{M}, x_0)$, and $[\beta]$ be its abelianization in $H_1(\mathbf{M},\mathbb{Q})$.
The function returns an \emph{intersection number} between two elements in $H_1(\mathbf{M},\mathbb Q)$. Assuming that $[\beta]$ is the second argument of i, a positive crossing occurs when $[\beta]$'s angle of intersection with the first homology class is positive. Fix a point $x_0\in\mathbf{M}$. $x_0$ lifts to $\tilde{x}_0\in\tilde{\mathbf{U}}$ by the covering map $f$, such that $\tilde{x}_0$ belongs to the quotient of $\tilde{\mathbf{U}}$ indexed $\mathbf{T}^{0,0}$.

$\tilde{\mathbf{U}}$ is associated to the kernel of the group homomorphism,
\begin{align}
\Omega_{u,v}:\pi_1(\mathbf{M},x_0)\rightarrow \mathbb{Z}^2 \text{; } \beta\mapsto(i(u,[\beta]),i(v,[\beta])).
\end{align}
Since $\mathbf{T}^{m,n}$ is isomorphic to $\mathbb{Z}^2$, the lift of $\beta\in \text{Ker } \Omega_{u,v}$ is the closed path, $\tilde{\beta}$. Let $F^t_\theta$ be a \emph{non-singular} geodesic flow on $\mathbf{M}\backslash$Sing$(\mathbf M)$ with initial point $x_0$. By [some theorem], $F^t_\theta$ is closed for for every rational direction $\theta$. The image of $F^t_\theta$ is an element of $\pi_1(\mathbf M, x_0)$, and homotopic to closed path $\beta$.

\begin{Def} Identify $\mathbb C$ with $\mathbb{R}^2$ in the usual way. The closed path $\alpha\in\pi_1(\mathbf M, x_0)$ is the path contained in the maximal cylinder decomposition in direction $(1,1)$.
\end{Def}

\begin{figure}[H]
\centering
\includesvg[pretex=\tiny, width=4.5in]{slopeonecylinder}
\caption{In the slope one direction, $\mathbf M$ decomposes into two cylinders.}
\end{figure}

We can see that $hol(\alpha)=6+6i$ is the holonomy vector of the closed loop $\alpha$ contained in one of these cylinders. Call the unlabeled cylinder $C_1$(right), and the labeled one $C_2$(left).
